---
title: "How far is Uluru to the coast"
author: "Will Kent"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How far is Uluru from the coast?
Uluru, Ayers Rock or simply "the Rock" is a large sandstone monolith located in central Australia. It is thought to have formed approximately 550 million years ago. It stands 348 metres above the surrounding plain and is listed as a UNESCO World Heritage Site.

Uluru is a site of imporance to the local *Pitjantjatjara* and *Yankunyjatjara* Aboriginal people. Uluru, and the surrounding area, is of great spiritual and cultural significance. 

This post will use R to answer the questions what is the minimum distance to the coast from Uluru? And where is that point?

To start the following libraries will need to be loaded. 

```{r library, message = FALSE, warning = FALSE}
library(tmaptools)
library(rgeos)
library(geosphere)
library(rgdal)
```

## Geospatial Data
Data relating to the surface of the earth, or geospatial data, can be a little confusing for the uninitiated so we'll start with some basics.

### Geospatial Objects
The basic geospatial objects are:

* Point - a physical location on the earth expressed by two coordinates (e.g.latitude and longitude),
* Linestring - a sequence of 2 or more connected points,
* Polygon - a closed collection of connected points that define a geographic boundary,
* Multipolygons - a collection of polygons.

### World Geodetic System
In the 1950's the United States Department of Defence, together with scientists from around the world, commenced efforts to establish a system to which geodetic data (measurements of the earth) could be referred. This collaboration established a common latitude, longitude and elevation co-ordinate system which led to the World Geodetic System 1960 (WGS60) standard.

Since 1960 the WGS has been updated a number of times to account for new information and data that effects measurements of the earths surface. The last such updated was in 1984 and formed the WGS84 used today as the standard reference system for modern technology based on the Gloabl Positioning Service (GPS). 

The use of WGS84 to place objects on the earths surface is a common practice but there are other methods available. In this instance WGS84 will be used as the co-ordinate reference system (CRS) and is set below for future use.
```{r projections}
wgs84 <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
```

## Locating Uluru
The first step in this process is to identify the latitude and logitude of Uluru using the WGS84 standard. There are a number of packages and services that provide the ability to obtain the latitude and longitude of an address or significant place but in this example the geocode_OSM function in the `tmaptools` package will be used. This function uses the Open Street Maps API to retrive the location of a place.
```{r uluru_geocode, message = FALSE, warning = FALSE}
loc <- geocode_OSM("Uluru", projection = wgs84)
longitude <- loc$coords["x"]
latitude <- loc$coords["y"]
```

The final step for the latitude and longitude that we've obtained is to convert the point to well known text (WKT). Well known text is a binary represenation ofa geospatial object and is needed for <insert reason>. The paste operation concatenates our latitude and longitude together in a string prefixed with _**POINT**_. The ensures the readWKT function from the `rgeos` is able to understand it is dealing with a point rather than another geospatial object.
```{r uluru_wkt, message = FALSE, warning = FALSE}
loc_txt <- paste("POINT(",longitude," ",latitude,")")
loc_wkt <- readWKT(loc_txt, p4s = CRS(wgs84))
```

## Coast line data
To help with the calculation of how far Uluru is from the coast fortunately there is a shape file of the coastlines of the world on the <http://www.naturalearthdata.com> website. In this step the shape file is downloaded locally and unzipped. The shape file is loaded into memory using the readOGR function in the `<packageName>` package.
```{r download, message = FALSE, warning = FALSE}
download.file("http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/physical/ne_10m_coastline.zip", 
              destfile = 'coastlines.zip')

unzip(zipfile = "coastlines.zip", 
      exdir = 'ne-coastlines-10m')

coast <- readOGR("ne-coastlines-10m/ne_10m_coastline.shp")
```

## Distance calculations
Now that the point location of Uluru and the coast lines of the world are known and loaded into R the minimum distance from Uluru to the coast can be calculated. This operation can be performed through the gDistance function in the `rgeos` package. The gDistance function finds the distance between two points, or in the case of a point and multiploygon, the minimum distance between the point and the polygon. 
```{r wgs84_distance, message = FALSE, warning = FALSE}
dist_line <- dist2Line(loc_wkt, coast)
closest_longitude <- dist_line[2]
closest_latitude <- dist_line[3]
```


```{r plot}
th     <- seq(0,2*pi,len=1000)
circle <- cbind(1.00001*min_degrees*cos(th)+loc_wkt$x,1.00001*min_degrees*sin(th)+loc_wkt$y)
sp_circle <- SpatialLines(list(Lines(list(Line(circle)),ID="1")),proj4string=CRS(wgs84))

# Plot on map
plot(coast, asp=1)
plot(loc_wkt, add = TRUE, col = "red", pch = 20)
plot(sp_circle, add = TRUE, col = "blue", lty = 2)
```
